---
title: "Extracting FLAG-tagged protein"
author: "Leonard Blaschek"
date: '2022-12-15'
output: 
  pdf_document:
    latex_engine: lualatex
    fig_caption: yes
    fig_height: 6
    includes:
      in_header: "preamble.tex"
mainfont: IBM Plex Sans
sansfont: IBM Plex Sans
monofont: Inconsolata
fontsize: 11 pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.kable.NA = '')

library(tidyverse)
library(kableExtra)
```


## Grinding the plant material

Grinding can be done in advance or on the day of extraction. Use force when grinding, and be thorough. No flakes of intact material should remain. Keep the powder in 50 ml falcons. When grinding in advance, store powder at -70°C until extraction.

## The day before -- prepare buffers

Washing and extraction buffers are identical, except for the addition of Triton X-100 to the latter. Calculate roughly 30 ml of each buffer per 5 g plant material, although it doesn't hurt to make some extra. For 100 ml each of washing and extraction buffer, start by preparing 200 ml of washing buffer:

```{r, echo=FALSE}
# Change the '200' into whichever final volume (extraction + washing buffer) you want to prepare
`Total volume (ml)` <- 200

# Only change code below if the concentrations of stock solutions have changed
dilution <- tribble(
  ~Component, ~`Stock conc. (M)`, ~`Final conc. (mM)`, ~`Volume (ml)`,
  "Tris-HCl, pH 7.4", 1, 20, NA,
  "NaCl", 4, 150, NA,
  "MgCl\\textsubscript{2}", 1, 5, NA,
  "EDTA", 0.5, 1, NA,
  "Water", NA, NA, NA
) %>%
  mutate(
    `Volume (ml)` = case_when(
    Component != "Water" ~
      `Final conc. (mM)` / (`Stock conc. (M)` * 1000) * `Total volume (ml)`
    ),
    `Volume (ml)` = case_when(
    Component == "Water" ~
      `Total volume (ml)` - sum(`Volume (ml)`, na.rm = TRUE),
    TRUE ~ `Volume (ml)`
    )
  ) %>% 
  add_row(Component = "Total", `Volume (ml)` = `Total volume (ml)`)

kable(dilution,
  "latex",
  caption = "Washing buffer composition.",
  escape = F,
  booktabs = T,
  linesep = ""
) %>%
  kable_styling(latex_options = c(
    "striped",
    "HOLD_position"
  )) %>% 
  row_spec(6, bold = T) %>%
  footnote(
    general = "Finally, add 4 protease inhibitor tablets (1 tablet/50 ml) and stir until dissolved",
    general_title = "",
    threeparttable = T,
    escape = F
  )
  
```

Then aliqout 99 ml of washing buffer and add 1 ml Triton X-100 to make the extraction buffer.

Lastly, prepare 1x PBS (10 ml per sample), by diluting the 10x PBS stock (D106) 1:9 with MilliQ.

\newpage

## Protein extraction

During the whole extraction process, **work on ice!**

- Add extraction buffer into the 50 ml falcons containing the ground material. Use a sample:buffer ration of 1:3 (*e.g.* 15 ml of buffer for 5 g of sample).
- Vortex thoroughly while the buffer is thawing the powder, until you have an even suspension without clumps.
- Agitate the tubes horizontally at 120 rpm in the cold room (4°C) for at least 30 min. In the meantime:
    - Cool down the Sorvall centrifuge in the chemical room (first one to the left) to 4°C, install the correct rotor (SS-34) and select the respective rotor code (05).
    - Aliquot 300 µl anti-FLAG (DYKDDDDK) beads per 5 g of sample (beads are stored in the D106 fridge) into an eppendorf tube. **Do not vortex or shake the beads!** Make sure you fully resuspend the beads before aliquoting to get the correct amount of beads per volume.
    - Equilibrate the beads by adding 1 ml of extraction buffer and incubating on ice for 3--5 min. Use the magnetic rack to remove the buffer, then repeat twice more. Finally resuspend in the starting volume of extraction buffer (*e.g.* 600 µl for 2 samples of 5 g each)
- Transfer the sample to the round-bottomed centrifuge tubes (in Ouda's office). Equilibrate their weight by adding extraction buffer if necessary.
- Centrifuge at 12 000 rpm and 4°C for 30 min.
- Carefully split the supernatant into three technical replicates (use the yellow screw-capped, flat-bottomed tubes). Usually, you can easily remove 5 ml per replicate (15 ml in total) without any carry-over. Discard the pellet.
    - Keep 30 µl of supernatant from each sample for Western blot QC.
- Add 100 µl of equilibrated anti-FLAG beads to each tube.
- Agitate tubes horizontally (120 rpm) at 4°C for 2 h.
- Use the magnetic rack to remove the supernatant -- The FLAG-tagged proteins are now adsorbed to the beads.
    - Keep 30 µl of supernatant per sample (10 µl per replicate, pooled) for Western blot QC, discard the rest.
- Add 500 ml of extraction buffer, then **with a cut pipette tip** transfer the buffer **including beads** to an eppendorf tube.
- Add another 500 ml of extraction buffer to the yellow-capped tube to carry over any remaining beads to the eppendorf.
- Use the magnetic rack to remove the supernatant.
- Wash the beads three more times in 1 ml extraction buffer. During each wash, remove the tubes from the magnetic rack and invert slowly to wash thoroughly.
- Wash the beads six times in 1 ml washing buffer. During each wash, remove the tubes from the magnetic rack and invert slowly to wash thoroughly.
- Wash the beads three times in 1x PBS. During each wash, remove the tubes from the magnetic rack and invert slowly to wash thoroughly.
    - Keep 30 µl per sample (10 µl per replicate, pooled) of beads suspended in PBS for Western blot QC before removing the supernatant of the last wash.
- Remove the supernatant, seal the tube with parafilm and store at -70°C until sending the sample to proteomics.

\newpage

## Western blot QC

Before sending the sample off for proteomics, check your extracts for tagged proteins. If you get no, or very weak, bands, troubleshoot vector design, infiltration and protein extraction as needed.

- Check that 1x TBST is prepared (100 ml 10x TBS + 899 ml H~2~O + 1 ml Tween20)
- Prepare a tube of 4x Laemmli buffer by adding 100 µl 1 M DTT (both in the freezer in D106).
- Add 10 µl 4x Laemmli buffer to each 30 µl sample and boil at 95°C for 5 min.
- Prepare a BioRad TGX mini gel in tris-glycine-SDS (TGX) running buffer.
    - When using only one gel, seal the opposite site of the cassette with two glass plates.
    - Assemble everything before loading to make sure the orientation of the gel cassette is correct.
- Load 10 µl of the Kaleidoscope ladder and the boiled samples.
- Run gel at 200 V for 30 min.
- Open a membrane transfer cassette and place the top half face-up on the bottom part of the transfer tray (**The membrane is on top -- do not touch!**).
- Break the gel out of its plastic casing using the metal spatula, cut away the wells.
- Carefully move the gel onto the membrane.
- Cover with the other half of the transfer cassette and clamp the tray tight.
- Run the 7 min transfer protocol ("1 midi/2 mini gels")
- Cut the membrane to the size of the gel -- Note down the orientation of the gel, **do not flip the membrane throughout the following steps.**
- Block unspecific protein adsorption with 5% milk in 1x TBST for 45 min under agitation.
- Incubate with 1:10 000 diluted primary FLAG antibodies in 5% milk/1x TBST over night at 4°C.
- Wash 3 times in 1x TBST for 15 min under agitation.
- Incubate in 1:10 000 diluted secondary antibody (rabbit-anti-mouse–HRP) in 5% milk/1x TBST (1.5 µl in 15 ml)
for 1 hour under agitation.
- Wash 3 times in 1x TBST for 10 min under agitation.
- Incubate for 1 min in a 1:1 mix of stable peroxide and substrate (5 ml total should suffice).
- Image on the black tray with the chemiluminescence program.
- Stain and de-stain with amido black (at least 10 min each) to visualise total protein.


