---
title: "Extracting FLAG-tagged protein"
author: "Leonard Blaschek"
date: '2022-12-15'
output: 
  pdf_document:
    latex_engine: lualatex
    fig_caption: yes
    fig_height: 6
mainfont: IBM Plex Sans
sansfont: IBM Plex Sans
monofont: Inconsolata
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.kable.NA = '')

library(tidyverse)
library(kableExtra)
```

## *Agrobacterium* infiltration of tobacco leaves

## Grinding the plant material

Grinding can be done in advance or on the day of extraction. Use force when grinding, and be thorough. No flakes of intact material should remain. Keep the powder in 50 ml falcons. When grinding in advance, store powder at -70°C until extraction.

## The day before -- prepare buffers

Washing and extraction buffers are identical, except for the addition of Triton X-100 to the latter. Calculate roughly 30 ml of each buffer per 5 g plant material, although it doesn't hurt to make some extra. For 100 ml each of washing and extraction buffer, start by preparing 200 ml of washing buffer:

```{r, echo=FALSE}
# Change the '200' into whichever final volume (extraction + washing buffer) you want to prepare
`Total volume (ml)` <- 200

# Only change code below if the concentrations of stock solutions have changed
dilution <- tribble(
  ~Component, ~`Stock conc. (M)`, ~`Final conc. (mM)`, ~`Volume (ml)`,
  "Tris-HCl, pH 7.4", 1, 20, NA,
  "NaCl", 4, 150, NA,
  "MgCl\\textsubscript{2}", 1, 5, NA,
  "EDTA", 0.5, 1, NA,
  "Water", NA, NA, NA
) %>%
  mutate(
    `Volume (ml)` = case_when(
    Component != "Water" ~
      `Final conc. (mM)` / (`Stock conc. (M)` * 1000) * `Total volume (ml)`
    ),
    `Volume (ml)` = case_when(
    Component == "Water" ~
      `Total volume (ml)` - sum(`Volume (ml)`, na.rm = TRUE),
    TRUE ~ `Volume (ml)`
    )
  ) %>% 
  add_row(Component = "Total", `Volume (ml)` = `Total volume (ml)`)

kable(dilution,
  "latex",
  caption = "Washing buffer composition.",
  escape = F,
  booktabs = T,
  linesep = ""
) %>%
  kable_styling(latex_options = c(
    "striped",
    "HOLD_position"
  )) %>% 
  row_spec(6, bold = T) %>%
  footnote(
    general = "Lastly, add 4 protease inhibitor tablets (1 tablet/50 ml)",
    general_title = "",
    threeparttable = T,
    escape = F
  )
  
```

Then aliqout 99 ml of washing buffer and add 1 ml Triton X-100 to make the extraction buffer.

Lastly, prepare 1x PBS (3 ml per sample), by diluting the 10x PBS stock (D106) 1:9 with MilliQ.

\newpage

## Protein extraction

During the whole extraction process, *work on ice!*

- Add extraction buffer into the 50 ml falcons containing the ground material. Use a sample:buffer ration of 1:3 (*e.g.* 15 ml of buffer for 5 g of sample).
- Vortex thoroughly while the buffer is thawing the powder, until you have an even suspension without clumps.
- Agitate the tubes horizontally at 12 rpm in the cold room (4°C) for at least 30 min. In the meantime:
    - Cool down the big centrifuge in the chemical room (first one to the left) to 4°C, install the correct rotor and select the respective rotor code.
    - Aliquot 100 µl anti-FLAG (DYKDDDDK) beads per 5 g of sample (beads are stored in the D106 fridge) into an eppendorf tube. *Do not vortex or shake the beads!* Make sure you fully resuspend the beads before aliquoting to get the correct amount of beads per volume.
    - Equilibrate the beads by adding 1 ml of extraction buffer and incubating on ice for 3--5 min. Use the magnetic rack to remove the buffer, then repeat twice more.
- Transfer the sample to the round-bottomed centrifuge tubes (in Ouda's office). Equilibrate their weight by adding extraction buffer if necessary.
- Centrifuge at 12 000 rpm and 4°C for 30 min.
- Carefully split the supernatant into three technical replicates (use the yellow screw-capped, flat-bottomed tubes). Usually, you can easily remove 5 ml per replicate (15 ml in total) without any carry-over. Discard the pellet.
    - Keep 50 µl of supernatant from each sample for Western blot QC.
- Add 100 µl of equilibrated anti-FLAG beads to each tube.
- Agitate tubes horizontally (120 rpm) at 4°C for 2 h.
- Use the magnetic rack to remove the supernatant -- The FLAG-tagged protein are now adsorbed to the beads.
    - Keep 50 µl of supernatant from each sample for Western blot QC, discard the rest.
- Add 500 ml of extraction buffer, then *with a cut pipette tip* transfer the buffer *including beads* to an eppendorf tube.
- Add another 500 ml of extraction buffer to the yellow-capped tube to carry over any remaining beads to the eppendorf.
- Use the magnetic rack to remove the supernatant.
- Wash the beads three more times in 1 ml extraction buffer. During each wash, remove the tubes from the magnetic rack and invert slowly to wash thoroughly.
- Wash the beads six times in 1 ml washing buffer. During each wash, remove the tubes from the magnetic rack and invert slowly to wash thoroughly.
- Wash the beads three times in 1x PBS. During each wash, remove the tubes from the magnetic rack and invert slowly to wash thoroughly.
    - Keep 50 µl of beads suspended in PBS for Western blot QC before removing the supernatant of the last wash.
- Remove the supernatant, seal the tube with parafilm and store at -70°C until sending the sample to proteomics.

\newpage

## Western blot QC

