---
title: "Sulis et al. (2023) re-analysis"
author: "Leonard Blaschek"
format:
  html:
    embed-resources: true
    theme: theme/sun_IBM.scss
    css: theme/styles.css
comments:
  hypothesis: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(kableExtra)
library(ggthemes)
library(showtext)
library(scales)
library(lubridate)
library(tidyverse)
library(ggrepel)
library(ggtext)

options(knitr.kable.NA = "")

#### import IBMPlexSans ####
font_add(
  "IBMPlexSans",
  regular = "/OFL_fonts/IBMPlexSans-Light.otf",
  italic = "/OFL_fonts/IBMPlexSans-LightItalic.otf",
  bold = "/OFL_fonts/IBMPlexSans-SemiBold.otf",
  bolditalic = "/OFL_fonts/IBMPlexSans-SemiBoldItalic.otf"
)
font_add(
  "IBMPlexMono",
  regular = "/OFL_fonts/IBMPlexMono-Regular.otf",
  italic = "/OFL_fonts/IBMPlexMono-Italic.otf",
  bold = "/OFL_fonts/IBMPlexMono-SemiBold.otf",
  bolditalic = "/OFL_fonts/IBMPlexMono-SemiBoldItalic.otf"
)
font_add(
  "Futura",
  regular = "/usr/share/fonts/propietary_fonts/Futura/FuturaStd-Medium.otf",
  italic = "/usr/share/fonts/propietary_fonts/Futura/FuturaStd-MediumOblique.otf",
  bold = "/usr/share/fonts/propietary_fonts/Futura/FuturaStd-Bold.otf",
  bolditalic = "/usr/share/fonts/propietary_fonts/Futura/FuturaStd-BoldOblique.otf"
)
showtext_auto()

#### generating plot theme ####
theme_leo <- function(base_size = 20,
                      base_family = "Futura") {
  theme_minimal(
    base_size = base_size,
    base_family = base_family
  ) %+replace%
    theme(
      strip.text = element_text(hjust = 0, face = "italic"),
      #       axis.ticks = element_line(
      #         linewidth = 0.25,
      #         lineend = "square",
      #         color = "black"
      #       ),
      axis.ticks = element_blank(),
      axis.text.x = element_text(
        colour = "black", # flipped coords
        margin = margin(1, 1, 1, 1),
        size = base_size
      ),
      axis.text.y = element_text(
        colour = "black",
        angle = 0,
        vjust = 0.5,
        hjust = 1,
        margin = margin(1, 1, 1, 1),
        size = base_size
      ),
      axis.title = element_text(size = base_size),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.line = element_line(lineend = "square"),
      # panel.border = element_rect(fill = NA, color = "black", size = 0.25),
      panel.spacing = unit(1.5, "mm"),
      legend.position = "bottom",
      legend.text = element_text(size = rel(0.8)),
      legend.key.height = unit(4, "mm"),
      complete = TRUE
    )
}

ggtext_size <- 20 / (14 / 5)
```

Most of us share the instinct to question our findings when presenting them to others. 
To highlight little quirks in the data that you cannot quite explain. 
As far as my experience goes, this instinct is strongest early on. 
*Don't immediately assume your data is just weird artifacts* is among my most commonly given advice when working with Bachelor or Master students. 
With time, we get a feeling what data we can trust (not a lot) and, when confronted with incisive questioning, might even get a little defensive.
There is, however, one form of communication that seems to make almost all of us walk back on our instincts and hesitate to share uncertainty: publishing.
Figures get adjusted until it shows what needs showing, but not much more.
Quirks, openly shared in the lab, get excluded from the manuscript so as to not invite unwelcome *questions.*
Why?
Because these *questions* will delay publication. 
They might result in calls for additional experiments or even rejection.
And that is to be avoided — we need that paper published; wee need it for our next grant, the next job, the chance to keep doing this silly job we really like.

# Annotated scatter plot

## Load Data

```{r}
volume_data <- read_tsv("/home/leonard/Dropbox/Reviewing/post publication review/2023_Sulis/volume.tsv")
height_data <- read_tsv("/home/leonard/Dropbox/Reviewing/post publication review/2023_Sulis/lignin.tsv")

data <- full_join(volume_data, height_data, by = "ID") |>
  mutate(label = case_when(
    str_detect(ID, "H-4|K-6|K-13|K-9|I-18|J-25") ~ ID,
    TRUE ~ NA
  ))

workshop_bar_data <- data |> 
  filter(!(is.na(label)) | str_detect(ID, "^WT4")) |> 
  mutate(ID = str_replace(ID, "WT", "WT-")) |> 
  select(
    ID, 
    "lignin" = `Total lignin*`, 
    "CL" = `C/L ratio`
    ) |> 
  separate(ID, into = c("strategy", "line", "replicate")) |> 
  unite("line", strategy, line, sep = "-") |> 
  mutate(line = case_when(line == "WT-4" ~ "Wildtype",
                          TRUE ~ line))

# write_tsv(workshop_bar_data, "Sulis2023_fig3EF.tsv")

sulis_scatter_data <- data |> 
  filter(!(str_detect(ID, "verage"))) |> 
  mutate(ID = str_replace(ID, "WT", "WT-")) |>
  mutate(type = case_when(str_detect(ID, "WT") ~ "Wildtype",
                           TRUE ~ "CRISPR-edited lines")) |> 
  select(
    ID, 
    type,
    label,
    "rel_lignin" = rel_lig, 
    rel_CL, 
    "rel_volume" = rel_vol) |> 
  separate(ID, into = c("strategy", "line", "replicate")) |> 
  unite("line", strategy, line, sep = "-") |> 
  mutate(line = case_when(str_detect(line, "WT") ~ "WT",
                          TRUE ~ line))

# write_tsv(workshop_scatter_data, "Sulis2023_fig3G.tsv")
```

## Plot lines

```{r}
panel_K <- ggplot(
  data = sulis_scatter_data,
  aes(
    x = rel_lignin,
    y = rel_volume
  )
) +
  geom_hline(
    yintercept = 100,
    colour = "grey",
    linetype = "dashed"
  ) +
  geom_point(
    aes(
      colour = type,
      fill = after_scale(colorspace::lighten(colour, 0.6))
    ),
    shape = "circle filled"
  ) +
  geom_label_repel(
    aes(label = label),
    size = ggtext_size,
    label.size = NA,
    fill = rgb(1, 1, 1, 0.5),
    min.segment.length = 0,
    family = "Futura"
  ) +
  scale_colour_manual(
    values = c("#275d95", "#d25952")
  ) +
  labs(
    x = "**Lignin content** (% of wildtype)",
    y = "<b>Stem volume</b> (% of wildtype)"
  ) +
  theme_leo() +
  theme(
    axis.title.x = element_markdown(),
    axis.title.y = element_markdown(),
    legend.position = c(0.15, 0.95),
    legend.title = element_blank()
  )


panel_K
# ggsave(
#   "/home/leonard/Dropbox/Reviewing/post publication review/2023_Sulis/sulis_highlight_plot.pdf",
#   width = 100,
#   height = 80,
#   units = "mm"
# )
```

# Annotated scatter plot with context

## Load Data

```{r}
volume_data <- read_tsv("/home/leonard/Dropbox/Reviewing/post publication review/2023_Sulis/volume.tsv")
height_data <- read_tsv("/home/leonard/Dropbox/Reviewing/post publication review/2023_Sulis/lignin.tsv")

data <- full_join(volume_data, height_data, by = "ID") |>
  mutate(label = case_when(
    str_detect(ID, "H-4|K-6") ~ ID,
    TRUE ~ NA
  ))

workshop_bar_data <- data |> 
  filter(!(is.na(label)) | str_detect(ID, "^WT4")) |> 
  mutate(ID = str_replace(ID, "WT", "WT-")) |> 
  select(
    ID, 
    "lignin" = `Total lignin*`, 
    "CL" = `C/L ratio`
    ) |> 
  separate(ID, into = c("strategy", "line", "replicate")) |> 
  unite("line", strategy, line, sep = "-") |> 
  mutate(line = case_when(line == "WT-4" ~ "Wildtype",
                          TRUE ~ line))

write_tsv(workshop_bar_data, "Sulis2023_fig3EF.tsv")

workshop_scatter_data <- data |> 
  filter(!(str_detect(ID, "verage"))) |> 
  mutate(ID = str_replace(ID, "WT", "WT-")) |>
  mutate(type = case_when(str_detect(ID, "WT") ~ "Wildtype",
                           TRUE ~ "CRISPR-edited lines")) |> 
  select(
    ID, 
    type,
    label,
    "rel_lignin" = rel_lig, 
    rel_CL, 
    "rel_volume" = rel_vol) |> 
  separate(ID, into = c("strategy", "line", "replicate")) |> 
  unite("line", strategy, line, sep = "-") |> 
  mutate(line = case_when(str_detect(line, "WT") ~ "WT",
                          TRUE ~ line))

# write_tsv(workshop_scatter_data, "Sulis2023_fig3G.tsv")
```

## Plot lines

```{r}
sulis_scatter_data <- read_tsv("/home/leonard/Applications/R_scripts/2023_ggplot2_workshop/data/Sulis2023_fig3G.tsv")

lit_data <- read_tsv("/home/leonard/Dropbox/Reviewing/post publication review/2023_Sulis/lit.tsv") |> 
  mutate(vol = height * pi * (diameter/2)^2) |> 
  group_by(paper) |> 
  mutate(rel_volume = vol / vol[line == "WT"] * 100,
         rel_lignin = lignin / lignin[line == "WT"] * 100,
         label = paper,
         type = "Literature") |> 
  filter(line != "WT") |> 
  select(rel_lignin, rel_volume, type, label)

sulis_scatter_data <- bind_rows(sulis_scatter_data, lit_data)

panel_K <- ggplot(
  data = sulis_scatter_data,
  aes(
    x = rel_lignin,
    y = rel_volume
  )
) +
  geom_hline(
    yintercept = 100,
    colour = "grey",
    linetype = "dashed"
  ) +
  geom_point(
    aes(
      colour = type,
      shape = type,
      fill = after_scale(colorspace::lighten(colour, 0.6))
    ),
    # shape = "circle filled"
  ) +
  geom_label_repel(
    aes(label = label),
    size = ggtext_size,
    label.size = NA,
    fill = rgb(1, 1, 1, 0.5),
    min.segment.length = 0,
    family = "Futura"
  ) +
  scale_colour_manual(
    values = c("#275d95", "#e8c245", "#d25952")
  ) +
  scale_shape_manual(
    values = c("circle filled", "square filled", "circle filled")
    ) +
  labs(
    x = "**Lignin content** (% of wildtype)",
    y = "<b>Stem volume</b> (% of wildtype)"
  ) +
  theme_leo() +
  theme(
    axis.title.x = element_markdown(),
    axis.title.y = element_markdown(),
    legend.position = c(0.15, 0.95),
    legend.title = element_blank()
  )


panel_K
# ggsave(
#   "/home/leonard/Dropbox/Reviewing/post publication review/2023_Sulis/sulis_highlight_plot_context.pdf",
#   width = 100,
#   height = 80,
#   units = "mm"
# )
```

# Model evaluation

## Load data

```{r}
actual <- data |>
  separate(ID, into = c("code", "line", "rep")) |>
  select(
    "code",
    "line",
    "rep",
    "Lignin content" = rel_lig,
    "Total sugar" = 28,
    "C/L ratio" = rel_CL,
    "Height" = rel_height,
    "Diameter" = rel_diam,
    "Volume" = rel_vol
  ) |> 
  pivot_longer(-c(code, line, rep), names_to = "variable", values_to = "actual")

pred <- read_tsv("/home/leonard/Dropbox/Reviewing/post publication review/2023_Sulis/predictions.tsv")
codes <- read_tsv("/home/leonard/Dropbox/Reviewing/post publication review/2023_Sulis/strategies_codes.tsv")
numbers <- read_tsv("/home/leonard/Dropbox/Reviewing/post publication review/2023_Sulis/strategies_number.tsv")

selected_lines <- numbers |>
  left_join(codes) |>
  filter(!(is.na(code))) |>
  select(Strategies, code)

eval_data <- pred |>
  right_join(selected_lines) |>
  select(code, Strategies, c(5:29)) |>
  pivot_longer(-c(code, Strategies), names_to = "variable", values_to = "predicted") |> 
  right_join(actual, by = c("code", "variable")) |> 
  drop_na(predicted) |> 
  group_by(code, variable) |> 
  summarise(
    actual = median(actual, na.rm = T),
    predicted = unique(predicted))

workshop_pred_data <- pred |> 
  select("strategies" = Strategies,
         "target_genes" = `Number of targeting genes`,
         "rel_lignin" = `Lignin content`,
         "rel_CL" = `C/L ratio`,
         "rel_volume" = Volume)

# write_tsv(workshop_pred_data, "Sulis2023_predictions.tsv")

```

## Plot model performance

```{r}
under50 <- tibble(actual = c(1:300)) |> 
  mutate(predicted = actual * 0.5,
         group = "under") |> 
  add_row(actual = 300, predicted = 0, group = "under")

over50 <- tibble(actual = c(1:300)) |> 
  mutate(predicted = actual * 2,
         group = "over") |> 
  add_row(actual = 0, predicted = 300, group = "over")

off50 <- bind_rows(under50, over50) 

eval_plot <- ggplot(
  eval_data,
  aes(
    x = actual,
    y = predicted
  )
) +
  geom_polygon(
    data = off50,
    aes(fill = group, group = group)
  ) +
  annotate(
    "text",
    label = "< 50% ",
    x = 280,
    y = 30,
    colour = "white",
    size = ggtext_size,
    hjust = 1,
    fontface = "italic",
    family = "Futura"
  ) +
  annotate(
    "text",
    label = "> 200% ",
    x = 30,
    y = 280,
    angle = 90,
    colour = "white",
    size = ggtext_size,
    hjust = 1,
    fontface = "italic",
    family = "Futura"
  ) +
  geom_abline(slope = 1, colour = "grey") +
  # geom_hline(yintercept = 100, colour = "grey") +
  # geom_vline(xintercept = 100, colour = "grey") +
  geom_point(
    colour = "black",
    fill = "white",
    shape = 21,
    size = 2,
    # stroke = 0.1
  ) +
  # geom_smooth(method = "lm") +
  labs(
    x = "Actual [% of WT]",
    y = "Predicted [% of WT]"
  ) +
  # geom_smooth(method = "lm") +
  scale_fill_manual(values = c(colorspace::desaturate("#275d95", 0.5), colorspace::desaturate("#d25952", 0.5))) +
  coord_fixed(xlim = c(20,280), ylim = c(20,280), expand = FALSE) +
  theme_leo() +
  theme(legend.position = "none") +
  facet_wrap(~variable)

eval_plot
# ggsave(
#   "/home/leonard/Dropbox/Reviewing/post publication review/2023_Sulis/sulis_eval_plot.pdf",
#   width = 120,
#   height = 120,
#   units = "mm"
# )
```
