---
title: "Hexokinase phosphorylation"
author: "Leonard Blaschek"
format: 
  html:
    embed-resources: true
execute:
  echo: false
  message: false
  warning: false
---

```{r}
library(tidyverse)
library(ggtext)
library(gt)
library(showtext)
library(tukeygrps)
library(patchwork)

showtext_opts(dpi = 400)

font_add(
  "Futura",
  regular = "FuturaStd-Medium.otf",
  italic = "FuturaStd-MediumOblique.otf",
  bold = "FuturaStd-Bold.otf",
  bolditalic = "FuturaStd-BoldOblique.otf"
)
showtext_auto()

text_size <- 6
ggtext_size <- text_size / (14 / 5)

#### generating plot theme ####
theme_leo <- function(base_size = text_size,
                      base_family = "Futura") {
  theme_minimal(
    base_size = base_size,
    base_family = base_family
  ) %+replace%
    theme(
      strip.text = element_text(hjust = 0, face = "italic"),
      axis.ticks = element_blank(),
      axis.text.x = element_text(
        colour = "black",
        margin = margin(1, 1, 1, 1),
        size = text_size
      ),
      axis.text.y = element_text(
        colour = "black",
        angle = 0,
        vjust = 0.5,
        hjust = 1,
        margin = margin(1, 1, 1, 1),
        size = text_size
      ),
      # axis.title = element_blank(),
      axis.line = element_line(
        linewidth = 0.4,
        colour = "black",
        lineend = "square"
      ),
      plot.title = element_text(size = text_size, hjust = 0),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.spacing = unit(1.5, "mm"),
      legend.position = "bottom",
      legend.text = element_text(size = rel(0.8)),
      legend.key.height = unit(4, "mm"),
      complete = TRUE
    )
}

pal_ostwald_disc <- c(
  "#275d95",
  "#e8c245",
  "#d25952"
)
```

# Note of caution

All presented values are the raw intensity of phosphorylated sites. 
Any increase can also simply reflect an increase in total protein at stable proportions of phosphorylation. 
An exact ratio of phosphorylated to unphosphorylated protein is hard to do here because of the FLAG/TiO enrichment steps.
If that seems crucial to you, we can try it nonetheless.

# Summary

S184 seems to get phosphorylated in prolonged darkness and after sucrose treatment. That doesn't seem intuitive to me, but what do I know. When I get some time, I'll see if I can estimate phosphorylated protein proportions (see *note of caution*).

# R1407

## Load DDA data

```{r}
dda <- read_tsv("/home/leonard/Dropbox/PUP-IT_blast/Delphine/2024/HXK_phospho/R1407/combined_protein.tsv") |>
  select(`Protein ID`, Gene, contains("MaxLFQ")) |>
  filter(`Protein ID` %in% c("Q42525", "P93834", "Q9LPS1")) |>
  pivot_longer(
    contains("MaxLFQ"),
    names_to = "sample",
    values_to = "protein_intensity"
  ) |>
  separate_wider_delim(
    sample,
    delim = "_",
    names = c("species", "construct", "bait", "treatment", "time", "replicate")
  ) |>
  mutate(
    time = ordered(time, levels = c("T0", "T8", "T16")),
    replicate = as.numeric(str_remove(replicate, " MaxLFQ Intensity"))
  )
```


## Load DIA-NN data

```{r}
manifest <- read_tsv(
  "/home/leonard/Dropbox/PUP-IT_blast/Delphine/2024/HXK_phospho/R1407/1407_TiO.fp-manifest",
  col_names = c("path", "sample", "replicate", "type"),
  show_col_types = FALSE
)

seqs <- read_tsv(
  "/home/leonard/Dropbox/PUP-IT_blast/Delphine/2024/HXK_phospho/R1407/HXK.tsv",
  col_names = c("Genes", "sequence"),
  show_col_types = FALSE
)

# ptm <- read_csv(
#   "/home/leonard/Dropbox/PUP-IT_blast/Delphine/2024/HXK_phospho/R1407/msstats_ptm.csv",
#   show_col_types = FALSE
# ) |>
#   filter(ProteinName %in% c("Q42525", "P93834", "Q9LPS1")) |>
#   filter(!is.na(`STY:79.96633`))

pr_report <- read_tsv(
  "/home/leonard/Dropbox/PUP-IT_blast/Delphine/2024/HXK_phospho/R1407/report.pr_matrix.tsv",
  show_col_types = FALSE
) |>
  filter(!is.na(`STY:79.96633`)) |>
  filter(`STY:79.96633 Best Localization` > 0.8) |>
  pivot_longer(starts_with("/data/"), names_to = "path", values_to = "intensity") |>
  left_join(manifest) |>
  separate_wider_delim(
    sample,
    delim = "_",
    names = c("species", "construct", "bait", "treatment", "time")
  ) |>
  filter(bait != "GFP") |>
  mutate(time = ordered(time, levels = c("T0", "T8", "T16")))

n_sites <- pr_report |>
  pull(Modified.Sequence) |>
  unique() |>
  length()

ggplot(
  pr_report,
  aes(y = log10(intensity), x = time)
) +
  geom_violin(
    fill = NA
  ) +
  geom_boxplot(
    colour = "black",
    width = 0.2,
    fatten = 1,
    outliers = FALSE
  ) +
  labs(
    title = paste("Distribution of", n_sites, "high-confidendence phosphosites"),
    y = "Log~10~(intensity)"
  ) +
  theme_leo() +
  theme(
    axis.title.y = element_markdown(),
    axis.title.x = element_blank(),
    plot.title = element_text(size = text_size * 1.5)
  )
```

## Summary of high-confidence HXK phosphosites

```{r}
hxk_report <- pr_report |>
  filter(Protein.Group %in% c("Q42525", "P93834", "Q9LPS1")) |>
  filter(bait == Genes) |>
  left_join(seqs) |>
  mutate("preceding" = nchar(str_extract(sequence, paste0("([:alpha:]+)", Stripped.Sequence), group = 1))) |>
  left_join(dda, by = join_by(
    "Protein.Group" == "Protein ID",
    "species",
    "construct",
    "bait",
    "treatment",
    "time",
    "replicate"
  )) |>
  mutate(intensity = (intensity / protein_intensity) * 100000)

hxk_summary <- hxk_report |>
  mutate(
    Sequence = paste0(
      "~",
      preceding + 1,
      "~",
      str_remove(
        str_replace(
          Modified.Sequence,
          "([:alpha:])\\(UniMod:21\\)",
          "**\\1**"
        ),
        "\\(UniMod:4\\)|\\(UniMod:35\\)"
      ),
      "~",
      preceding + nchar(Stripped.Sequence),
      "~"
    )
  ) |>
  group_by(Genes, Sequence, time) |>
  summarise(intensity = mean(intensity, na.rm = T)) |>
  pivot_wider(names_from = time, values_from = intensity) |>
  group_by(Genes)

gt(
  hxk_summary,
  row_group_as_column = TRUE
) |>
  fmt_markdown(columns = Sequence) |>
  fmt_number(
    columns = c(3, 4),
    decimals = 2
  ) |>
  # cols_width(
  # Sequence ~ pct(50)
  # ) |>
  cols_nanoplot(
    columns = starts_with("T"),
    new_col_name = "nanoplots",
    new_col_label = md("Progression (0–16 h)")
  )
```


## Plot sites around position 180

```{r}
focus_sites <- hxk_report |>
  filter(
    str_detect(Stripped.Sequence, "SLSSG")
    # | str_detect(Modified.Sequence, "ELGFTFS\\(")
  ) |>
  mutate(
    "Site" = case_when(
      str_detect(Modified.Sequence, "ELGFTFS\\(") ~ "p177",
      str_detect(Modified.Sequence, "SSGS\\(") ~ "p189",
      str_detect(Modified.Sequence, "SLSS\\(") ~ "p187",
      str_detect(Modified.Sequence, "SLS\\(") ~ "p186",
      str_detect(Modified.Sequence, "\\)LSS") ~ "p184",
      TRUE ~ NA
    ),
    bait = paste(bait, "bait")
  )

ggplot(
  focus_sites,
  aes(x = time, y = intensity, colour = Genes)
) +
  stat_summary(
    aes(group = Genes, colour = Genes),
    geom = "line",
    fun = "mean"
  ) +
  stat_summary(
    geom = "pointrange",
    fun.data = "mean_se",
    size = 0.3
  ) +
  facet_grid(bait ~ Site) +
  scale_colour_manual(
    values = pal_ostwald_disc[c(1, 3)],
    aesthetics = c("colour", "fill"),
    name = "Protein"
  ) +
  labs(
    y = "Rel. phosphorylated proportion",
  ) +
  theme_leo() +
  theme(
    axis.title.x = element_blank(),
    axis.text.y = element_blank()
    )
```

# R1151

The only peptide for 177 with high-confidence phosphosite localisation found in this experiment is identical between HXK1 and HXK2.
That means we cannot distinguish between HXK1 and HXK2 for that site, explaining the counterintuitive intensity values in the figures below.

## Load DDA data

```{r}
dda <- read_tsv("/home/leonard/Dropbox/PUP-IT_blast/Delphine/2024/HXK_phospho/R1151/combined_protein.tsv") |>
  select(`Protein ID`, Gene, contains("MaxLFQ")) |>
  filter(`Protein ID` %in% c("Q42525", "P93834", "Q9LPS1")) |>
  pivot_longer(
    contains("MaxLFQ"),
    names_to = "sample",
    values_to = "protein_intensity"
  ) |>
  separate_wider_delim(
    sample,
    delim = "_",
    names = c("species", "construct", "bait", "treatment", "time", "replicate")
  ) |>
  mutate(
    time = ordered(time, levels = c("4h", "24h")),
    treatment = case_when(
      treatment == "sucrose" ~ "suc.",
      treatment == "mannitol" ~ "man."
    ),
    replicate = as.numeric(str_remove(replicate, " MaxLFQ Intensity"))
  )
```


## Load DIA-NN data

```{r}
manifest <- read_tsv(
  "/home/leonard/Dropbox/PUP-IT_blast/Delphine/2024/HXK_phospho/R1151/R1151_TiO.fp-manifest",
  col_names = c("path", "sample", "replicate", "type"),
  show_col_types = FALSE
)

seqs <- read_tsv(
  "/home/leonard/Dropbox/PUP-IT_blast/Delphine/2024/HXK_phospho/R1407/HXK.tsv",
  col_names = c("Genes", "sequence"),
  show_col_types = FALSE
)

# ptm <- read_csv(
#   "/home/leonard/Dropbox/PUP-IT_blast/Delphine/2024/HXK_phospho/R1407/msstats_ptm.csv",
#   show_col_types = FALSE
# ) |>
#   filter(ProteinName %in% c("Q42525", "P93834", "Q9LPS1")) |>
#   filter(!is.na(`STY:79.96633`))

pr_report <- read_tsv(
  "/home/leonard/Dropbox/PUP-IT_blast/Delphine/2024/HXK_phospho/R1151/report.pr_matrix.tsv",
  show_col_types = FALSE
) |>
  filter(!is.na(`STY:79.96633`)) |>
  filter(`STY:79.96633 Best Localization` > 0.8) |>
  pivot_longer(starts_with("/data/"), names_to = "path", values_to = "intensity") |>
  left_join(manifest) |>
  separate_wider_delim(
    sample,
    delim = "_",
    names = c("species", "construct", "bait", "treatment", "time")
  ) |>
  filter(bait != "GFP") |>
  mutate(
    time = ordered(time, levels = c("4h", "24h")),
    treatment = case_when(
      treatment == "sucrose" ~ "suc.",
      treatment == "mannitol" ~ "man."
    )
  )

n_sites <- pr_report |>
  pull(Modified.Sequence) |>
  unique() |>
  length()

ggplot(
  pr_report,
  aes(y = log10(intensity), x = bait)
) +
  geom_violin(
    fill = NA
  ) +
  geom_boxplot(
    colour = "black",
    width = 0.2,
    fatten = 1,
    outliers = FALSE
  ) +
  labs(
    title = paste("Distribution of", n_sites, "high-confidendence phosphosites"),
    y = "Log~10~(intensity)"
  ) +
  theme_leo() +
  theme(
    axis.title.y = element_markdown(),
    axis.title.x = element_blank(),
    plot.title = element_text(size = text_size * 1.5)
  )
```

## Summary of high-confidence HXK phospho-peptides

```{r}
hxk_report <- pr_report |>
  filter(Protein.Group %in% c("Q42525", "P93834", "Q9LPS1")) |>
  filter(bait == Genes) |>
  left_join(seqs) |>
  mutate("preceding" = nchar(str_extract(sequence, paste0("([:alpha:]+)", Stripped.Sequence), group = 1))) |>
  left_join(dda, by = join_by(
    "Protein.Group" == "Protein ID",
    "species",
    "construct",
    "bait",
    "treatment",
    "time",
    "replicate"
  )) |>
  mutate(intensity = (intensity / protein_intensity) * 100000)

hxk_summary <- hxk_report |>
  mutate(
    Sequence = paste0(
      "~",
      preceding + 1,
      "~",
      str_remove(
        str_replace(
          Modified.Sequence,
          "([:alpha:])\\(UniMod:21\\)",
          "**\\1**"
        ),
        "\\(UniMod:4\\)|\\(UniMod:35\\)"
      ),
      "~",
      preceding + nchar(Stripped.Sequence),
      "~"
    )
  ) |>
  group_by(Genes, Sequence, time) |>
  summarise(
    "response" = mean(intensity[treatment == "suc."], na.rm = T) / mean(intensity[treatment == "man."], na.rm = T)
  ) |>
  pivot_wider(names_from = time, values_from = response) |>
  group_by(Genes) |>
  arrange(Genes, `4h`)

gt(
  hxk_summary,
  row_group_as_column = TRUE
) |>
  fmt_markdown() |>
  fmt_number(
    columns = c(3, 4),
    decimals = 2
  ) |>
  cols_width(
    Sequence ~ pct(60)
  ) |>
  tab_spanner(
    label = "Rel. sucrose response",
    columns = c(`4h`, `24h`)
  )
```


## Plot sites around position 180

```{r}
focus_sites <- hxk_report |>
  filter(
    str_detect(Stripped.Sequence, "SLSSG")
    # | str_detect(Modified.Sequence, "ELGFTFS\\(")
  ) |>
  mutate(
    "Site" = case_when(
      str_detect(Modified.Sequence, "ELGFTFS\\(") ~ "p177",
      str_detect(Modified.Sequence, "SSGS\\(") ~ "p189",
      str_detect(Modified.Sequence, "SLSS\\(") ~ "p187",
      str_detect(Modified.Sequence, "SLS\\(") ~ "p186",
      str_detect(Modified.Sequence, "\\)LSS") ~ "p184",
      TRUE ~ NA
    ),
    bait = paste(bait, "bait")
  ) |>
  group_by(Site, bait, Genes, time, treatment, replicate) |>
  summarise(intensity = sum(intensity))

ggplot(
  focus_sites |> filter(time == "4h"),
  aes(x = treatment, y = intensity, colour = Genes)
) +
  stat_summary(
    aes(group = Genes, colour = Genes),
    geom = "line",
    fun = "mean"
  ) +
  stat_summary(
    geom = "pointrange",
    fun.data = "mean_se",
    size = 0.3
  ) +
  facet_grid(bait ~ Site) +
  scale_colour_manual(
    values = pal_ostwald_disc[c(1, 3)],
    aesthetics = c("colour", "fill"),
    name = "Protein"
  ) +
  labs(
    y = "Rel. phosphorylated proportion",
    title = "4 hours"
  ) +
  theme_leo() +
  theme(
    axis.title.x = element_blank(),
    axis.text.y = element_blank()
  )
```


```{r}
ggplot(
  focus_sites |> filter(time == "24h"),
  aes(x = treatment, y = intensity, colour = Genes)
) +
  stat_summary(
    aes(group = Genes, colour = Genes),
    geom = "line",
    fun = "mean"
  ) +
  stat_summary(
    geom = "pointrange",
    fun.data = "mean_se",
    size = 0.3
  ) +
  facet_grid(bait ~ Site) +
  scale_colour_manual(
    values = pal_ostwald_disc[c(1, 3)],
    aesthetics = c("colour", "fill"),
    name = "Protein"
  ) +
  labs(
    y = "Rel. phosphorylated proportion",
    title = "24 hours"
  ) +
  theme_leo() +
  theme(
    axis.title.x = element_blank(),
    axis.text.y = element_blank()
  )
```
